### 2012-03-11 intention motor representation joint action (paper 3)/intention and motor representation in joint action.tex
just started first draft of paper version, not much done

\begin{abstract}
On the assumption that social motor representation plays a role in explaining how effective joint action is possible, do we also need motor representation to explain what joint action is?  Philosophers tend to assume that motor representation is only an enabling condition for joint action and of no direct interest to narrowly philosophical theories of joint action and shared intention.  In this talk I shall argue that social motor representation and shared intention  have distinctive roles in explaining the purposiveness of joint action.  This gives rise to a challenge.  On the one hand, effective joint action---imagine two people erecting a tent in a gale together---sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.  On the other hand, recognizing their distinctive roles precludes the existence of direct inferential links between shared intentions and social motor representations.  The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite the lack of inferential integration.
\end{abstract}

\section{Slide}
[hello]

\section{Slide}
As a philosopher I’m going to start from an empirically controversial premise.
I don’t think this premise has been established, but I do think it’s a reasonable bet.

\section{Slide}
A motor representation is \emph{agent-neutral} if it concerns an action which is not one’s own or, in the case of joint action, not entirely one’s own.

\section{Slide}
Two or more motor representations are \emph{reciprocal} just if there is a single outcome which each motor representation represents.

\section{Slide}
Sometimes I’ll using the term \emph{social} motor representation for reciprocal agent-neutral motor representations.
I have a slightly bad conscience about this because I’m not sure they’re actually social in any ordinary sense.

It is almost uncontroversial that agent-neutral motor representations exist, and I see no reason to doubt that they could be reciprocal.
The controversial part of the premise is the claim that such representations can enable joint action.

I should say, by the way, that I’m not suggesting that reciprocal agent-neutral motor representation is involved in every joint action.
Perhaps some joint actions do not involve social motor representation.
The premise is just that social motor representation is among the factors which enable some joint actions.

\section{Slide}
I’m not going to defend the premise, although I will try to explain the premise in more detail in a moment.  But first let me outline where I’m going with this.

\section{Slide}
The premise leads to a question and, relatedly, to a challenge.

\section{Slide}
The question is whether social motor representation is not only an enabling condition for joint action but also plays a role in explaining what joint action is. I want to argue that it does.  In fact I want to argue that it plays a role similar and complementary to that of shared intention.

\section{Slide}
After this, in the last part of my talk, I also want to point to a challenge raised by the existence of social motor representation.  
The challenge is roughly this: planning for joint action involves representations of target outcomes in two different formats, a motor format and a propositional format.  
Given that representations in these different formats cannot straightforwardly be inferentially integrated in practical reasoning, what could ever insure that there is sometimes non-accidental harmony between motor representations and shared intentions?

\section{Slide}
Let me start by explaining the premise: Reciprocal agent-neutral motor representation enables joint action.
Why think this?
A direct way to test the premise would be to selectively intervene on social motor representation and see how it affects agents’ performance of joint actions.
As far as I know no one has done this.
But there are some indirect findings.
I shall mention just one, a finding by Dimitris Kourtis and colleagues.

\section{Slide}
They found that `the social relation between individuals modulates action simulation' \citep[p.\ 1]{kourtis:2010_favoritism}.

\section{Slide}
In slightly more detail:
\begin{quote}
`motor activation during action anticipation depends on the ... relation between the actor and the observer ... Simulation of another person’s action, as reﬂected in the activation of motor cortices, gets stronger the more the other is perceived as an interaction partner.’  \citep[p.\ 4]{kourtis:2010_favoritism}
\end{quote}
I realise this doesn’t conclusively support the premise.  But I think it is evidence in its favour.

\section{Slide}
Here’s the experiment they did.
The task was simple: two people sat opposite each other.  Sometimes they acted alone, picking up and replacing an object.  And sometimes had to act together, passing an object between them.  Also present was the ‘loner’ who always acted alone.

\section{Slide}
EEG measurements of motor activation were recorded.

\section{Slide}
[image shrinks]

\section{Slide}
The researchers found that patterns of activation for self and joint action partner were similar, and different from patterns of activation for the actions of the loner which were similar to patterns in a ‘no go’ condition where no one was to act.

\section{Slide}
So here’s what I take from this paper.
If you are engaged in a joint action with someone, one which involves moving an object by passing it between you, then each of you has motor representations of the other’s actions and these motor representations are functionally equivalent to motor representations of your own actions in the sense that they are just the sorts of representation that might have caused you to do what the other is doing (if you were in her position).

\section{Slide}
Suppose that Kourtis et al are right that social motor representation is more likely to occur in joint action than when one is merely observing.
How could this facilitate joint action?

\section{Slide}
Let’s step back and consider an individual action.
An agent moves a mug from one place to another, passing in from her left hand to her right hand half way [*demonstrate].
It’s a familiar idea that motor representations for planning and monitoring action involve an hierarchical structure,
where there is a relatively abstract representation of an outcome that is progressively filled in.

>>>>

\begin{abstract}
\noindent
On the assumption that social motor representation plays a role in explaining how effective joint action is possible, is there also a role for motor representation in explaining what joint action is?  
Philosophers tend to assume that motor representation is only an enabling condition for joint action and of no direct interest to narrowly philosophical theories of joint action and shared intention.  In this talk I shall argue that social motor representation and shared intention  have distinctive roles in explaining the purposiveness of joint action.  This gives rise to a challenge.  On the one hand, effective joint action---imagine two people erecting a tent in a gale together---sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.  On the other hand, recognizing their distinctive roles precludes the existence of direct inferential links between shared intentions and social motor representations.  The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite the lack of inferential integration.
\end{abstract}

\section{Introduction}
Which events are joint actions?
We shall argue that fully answering this question involves appeal to structures of motor representation.
By contrast, 
most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a notion of shared (or `collective') intention.
For an event to be a joint action is for it to be appropriately related to a shared intention.
For comparison, one might also hold that for an event to be an (ordinary, individual) action is for it to be appropriately related to an intention.
This assumption immediately raises the question of what shared intention is.
On this question several philosophers offer competing theories.
It is striking, however, that few have challenged the assumption that for an event to be a joint action it must be appropriately related to a shared intention.
This is striking because the parallel assumption about ordinary, individual action is not as widely held among leading theorists.
On several leading views, there could be agents whose actions are purposive---and, in some cases, even intentional---even though the agents have no intentions.
If any such view is correct, it is impossible to fully explain what it is for an event to be an action by appeal to intention alone.
Our aim is to defend a parallel claim about joint action.
More specifically, we shall defend these claims: 
first, 
that fully explaining which events are joint actions requires appeal to structures of motor representation and not only to shared intention;
and, 
second, 
that recognising a constitutive role for motor representation in joint action makes possible an account of shared intention which is simpler and more elegant than any of the leading alternatives.

\section{Social motor representation}
We start from the premise that reciprocal agent-neutral motor representation sometimes enables joint action.
This needs unpacking.
A motor representation is \emph{agent-neutral} if it concerns an action which is not one’s own or, in the case of joint action, not entirely one’s own.
Two or more motor representations are \emph{reciprocal} just if there is a single outcome which each motor representation represents.
We shall use the term \emph{social} motor representation as an abbreviation for reciprocal agent-neutral motor representation.
(Of course we cannot assume in advance of argument that reciprocal agent-neutral motor representations are social in any interesting sense;
but we are using the term `social' in a non-standard way as an abbreviation, one that reflects our aim.)
It is almost uncontroversial that social motor representations exist.
The controversial part of the premise is the claim that such representations can enable joint action.

Our premise should not be interpreted as implying that social motor representation is involved in every joint action,
nor that social motor representation is sufficient for joint action.
The premise is just that social motor representation is among the factors enabling some joint actions.

Why accept the premise that social motor representation enables joint action?
A direct way to test the premise would be
measure how (if at all) selective interventions on social motor representation affected agents’ performance of joint actions.
As far as we know no one has done this and it is probably impossible.
But there are some indirect findings.
We shall focus on just one, a finding by Kourtis and colleagues that `Simulation of another person’s action, as reﬂected in the activation of motor cortices, gets stronger the more the other is perceived as an interaction partner’  \citep[p.\ 4]{kourtis:2010_favoritism}.
This finding falls well short of providing conclusively support for the premise; we mention it as an illustration of the sort of evidence is relevant to the premise.

Here’s the experiment they did.
The task was simple: two people sat opposite each other.  Sometimes they acted alone, picking up and replacing an object.  And sometimes had to act together, passing an object between them.  Also present was the ‘loner’ who always acted alone.
EEG measurements of motor activation were recorded.
The researchers found that patterns of activation for self and joint action partner were similar, and different from patterns of activation for the actions of the loner which were similar to patterns in a ‘no go’ condition where no one was to act.
Since the researchers did not include a condition where there was a competitive, non-joint interaction, we cannot conclude that the modulation in action simulation they found is specific to joint action as opposed to other kinds of interaction.
What we do take from this paper,
together with a body of other research on motor simulation of action,
is this.
If you are engaged in a joint action with someone, one which involves moving an object by passing it between you, then each of you has motor representations of the other’s actions and these motor representations are functionally equivalent to motor representations of your own actions in the sense that they are just the sorts of representation that might have caused you to do what the other is doing (if you were in her position).

\section{How could social motor representation enable joint action?}
Suppose that Kourtis et al are right that social motor representation is more likely to occur in joint action than when one is merely observing.
It doesn't follow, of course, that social motor representation enables some joint actions (and nor do these researchers claim that it does).
But let us consider just the possibility that it might.
How could social motor representation enable joint action even in principle?

To answer this question
let us take a step back and consider an individual action.
Suppose an agent moves a mug from one place to another, passing in from her left hand to her right hand half way.
It is a familiar idea that motor representations for planning and monitoring action involve an hierarchical structure,
where there is a relatively abstract representation of an outcome that is progressively filled in.
